## Confidence interval exercises


## Exercises 

For these exercises, we will use actual polls from the 2016 election. You can load the data from the __dslabs__ package.

```{r}
library(dslabs)
data("polls_us_election_2016")
```

Specifically, we will use all the national polls that ended within one week before the election.

```{r, message=FALSE, message=FALSE}
library(tidyverse)
polls <- polls_us_election_2016 |> 
  filter(enddate >= "2016-10-31" & state == "U.S.") 
```

1\. For the first poll, you can obtain the samples size and estimated Clinton percentage with:
  
```{r, eval=FALSE}
N <- polls$samplesize[1]
x_hat <- polls$rawpoll_clinton[1]/100
```

Assume there are only two candidates and construct a 95% confidence interval for the election night proportion $p$. 
```{r}
library(tidyverse)
library(dslabs)
data("polls_us_election_2016")
polls <- polls_us_election_2016 %>% 
  filter(enddate >= "2016-10-31" & state == "U.S.") 
N <- polls$samplesize[1]
x_hat <- polls$rawpoll_clinton[1]/100
se_hat <- sqrt(x_hat * (1 - x_hat) / N)
c(x_hat - 1.96 * se_hat, x_hat + 1.96 * se_hat)
```


2\. Now use `dplyr` to add a confidence interval as two columns, call them `lower` and `upper`, to the object `poll`. Then use `select` to show the `pollster`, `enddate`, `x_hat`,`lower`, `upper` variables. Hint: define temporary columns `x_hat` and `se_hat`. 
```{r}
poll <- polls |> mutate(x_hat2 = rawpoll_clinton/100, 
                        se_hat = sqrt(x_hat2 * (1- x_hat2) / samplesize), 
                        lower = x_hat2 - 1.96 *se_hat,
                        upper = x_hat2 + 1.96 * se_hat) |>
  select(pollster, enddate, x_hat2, lower, upper)
head(poll)
```


3\. The final tally for the popular vote was Clinton 48.2%	and Trump 46.1%. Add a column, call it `hit`, to the previous table stating if the confidence interval included the true proportion $p=0.482$ or not. 
```{r}
p <- 0.482
poll <- poll |> mutate (hit = ifelse(lower < p & upper > p, 1, 0))
poll <- mutate(poll, p_inside = ifelse(hit, "Yes", "No"))
poll <- poll |> mutate(hit2 = lower <= 0.482 & upper >= 0.482)
head(poll)
```


4\. For the table you just created, what proportion of confidence intervals included $p$?
```{r}
mean(poll$hit)

```
 
  
5\. If these confidence intervals are constructed correctly, and the theory holds up, what proportion should include $p$?
polls may be bias, may be different in different poll
not the same because the its is not random sample.
```{r}
pnorm(1.96) - pnorm(-1.96)
```


```{r}
library(dslabs)
library(tidyverse)
data(heights)
x <- heights |> filter(sex == "Male") |> pull(height)
```

1. Mathematically speaking, x is our population. What are the average and standard deviation of our population?

```{r}
mean(x)
sd(x)
```


2. Call the population average computed above mu and the standard deviation sigma. Now take a sample of size 50, with replacement, and construct an estimate for mu and sigma. 

```{r}
set.seed(1)
N <- 50
X <- sample(x, N, replace = TRUE)
mean(X)
sd(X)
```

3. What does the theory tell us about the sample average xbar and how it is related to mu?
a. It is pratically identical to mu.
b. It is a random variable with expected value mu and standard error sigma/sqrt(N) -> correct. The sample average is a random variable whose standard error decreases with inreasing sample size. 
c. It is a random variable with expected value mu and standard error sigma.
d. It contains no information.

5. Based on the result construct a 95% confidence interval for mu using your sample from question 2
```{r}
c(mean(X) - 1.96 * sd(X), mean(X) + 1.96 * sd(X))
```

6. Now run a Monte Carlo simulation in which you compute 10,000 confidence intervals as you have just done. What proportion of these interval include mu?
```{r}
set.seed(1)
mu <- mean(x)
n <- 50
B <- 10000

res <- replicate(B, {
  X <- sample(x, N, replace = TRUE)
  interval <- mean(X) + c(-1,1)*1.96 * sd(X)/sqrt(N)
  between(mu, interval[1], interval[2])
})
mean(res)
```

  
6\. A much smaller proportion of the polls than expected produce confidence intervals containing $p$. If you look closely at the table, you will see that most polls that fail to include $p$ are underestimating. The reason for this is undecided voters, individuals polled that do not yet know who they will vote for or do not want to say. 
Because, historically, undecideds divide evenly between the two main candidates on election day, it is more informative to estimate the spread or the difference between the proportion of two candidates $d$, which in this election was $0. 482 - 0.461 = 0.021$. 
Assume that there are only two parties and that $d = 2p - 1$, redefine `polls` as below and
re-do exercise 1, but for the difference.

```{r, message=FALSE, comment=FALSE}
polls <- polls_us_election_2016 |> 
  filter(enddate >= "2016-10-31" & state == "U.S.")  |>
  mutate(d_hat = rawpoll_clinton / 100 - rawpoll_trump / 100)
N <- polls$samplesize[1]
d_hat <- polls$d_hat[1]
x_hat <- (d_hat+1) /2
se_hat <- 2 * sqrt(x_hat * (1 - x_hat) / N)
c(d_hat - 1.96 * se_hat, d_hat + 1.96 * se_hat)
```

7\. Now repeat exercise 3, but for the difference.
```{r}
library(tidyverse)
library(dslabs)
d<-0.021
data("polls_us_election_2016")
polls <- polls_us_election_2016 %>% 
  filter(enddate >= "2016-10-31" & state == "U.S.") %>%
  mutate(d_hat = rawpoll_clinton / 100 - rawpoll_trump / 100)
poll<-polls %>% select(pollster,enddate,samplesize,d_hat) 
poll<-poll %>% mutate(x_hat=(d_hat+1)/2)
poll<-poll %>% mutate(se_hat=2*sqrt(x_hat * (1 - x_hat) / samplesize))
poll<-poll %>% mutate(lower=d_hat - 1.96 * se_hat, upper=d_hat + 1.96 * se_hat)
poll<-poll %>% mutate(hit=ifelse(lower<= d & upper >= d,1,0)) %>% select(pollster,lower,upper,hit)
poll
```


8\. Now repeat exercise 4, but for the difference.
```{r}
mean(poll$hit)
```


9\. Although the proportion of confidence intervals goes up substantially, it is still lower than 0.95. In the next chapter, we learn the reason for this. To motivate this, make a plot of the error, the difference between each poll's estimate and the actual $d=0.021$. Stratify by pollster.
```{r}
library(tidyverse)
library(dslabs)
d<-0.021
data("polls_us_election_2016")
polls <- polls_us_election_2016 %>% 
  filter(enddate >= "2016-10-31" & state == "U.S.") %>%
  mutate(d_hat = rawpoll_clinton / 100 - rawpoll_trump / 100)
poll<-polls %>% select(pollster,enddate,samplesize,d_hat) 
poll<-poll %>% mutate(x_hat=(d_hat+1)/2)
poll<-poll %>% mutate(se_hat=2*sqrt(x_hat * (1 - x_hat) / samplesize))
poll<-poll %>% mutate(lower=d_hat - 1.96 * se_hat, upper=d_hat + 1.96 * se_hat)
poll<-poll %>% select(-samplesize)
poll<-poll %>% mutate(hit=ifelse(lower<= d & upper >= d,1,0)) %>% select(pollster,lower,upper,hit,d_hat) %>%  arrange(pollster) %>% mutate(pollnr=1:nrow(poll))
poll<-mutate(poll, d_inside = ifelse(hit, "Yes", "No") )
poll<-mutate(poll, pollsternr=paste(pollnr,pollster,sep="_"))
poll$d_error<-poll$d_hat-0.02
ggplot(poll, aes(pollster, d_error, ymin=lower, ymax=upper, col = d_inside)) + 
  geom_point()+
  #geom_errorbar() + 
  coord_flip() + 
  geom_hline(yintercept = 0)
```

10\. Redo the plot that you made for exercise 9, but only for pollsters that took five or more polls.

```{r}
library(tidyverse)
library(dslabs)
d<-0.021
data("polls_us_election_2016")
polls <- polls_us_election_2016 %>% 
  filter(enddate >= "2016-10-31" & state == "U.S.") %>%
  mutate(d_hat = rawpoll_clinton / 100 - rawpoll_trump / 100) #%>% 
  #mutate(pollnr=1:nrow(poll))
pollsters <- polls %>% select(pollster) %>% group_by(pollster) %>% summarise(n=n()) %>% arrange(desc(n)) %>% filter(n>=5)
polls<-left_join(pollsters,polls) %>% select(pollster,n,samplesize,rawpoll_clinton,rawpoll_trump,d_hat)
poll<-polls %>% select(pollster,samplesize,d_hat) 
poll<-poll %>% mutate(x_hat=(d_hat+1)/2)
poll<-poll %>% mutate(se_hat=2*sqrt(x_hat * (1 - x_hat) / samplesize))
poll<-poll %>% mutate(lower=d_hat - 1.96 * se_hat, upper=d_hat + 1.96 * se_hat)
poll<-poll %>% select(-samplesize)
poll<-poll %>% mutate(hit=ifelse(lower<= d & upper >= d,1,0)) %>% select(pollster,d_hat,lower,upper,hit) %>%  mutate(pollnr=1:nrow(poll))
poll <- mutate(poll, d_inside = ifelse(hit, "Yes", "No") )
poll$d_error<-poll$d_hat-0.02
ggplot(poll, aes(pollster, d_error, ymin=lower, ymax=upper, col = d_inside)) + 
  geom_point()+
  #geom_errorbar() + 
  coord_flip() + 
  geom_hline(yintercept = 0)
```


