---
title: "Exercises_Solutions_Parameters_SK"
author: "Soomin Kim"
date: "2022-10-15"
output: html_document
---


## Exercises

1\. Suppose you poll a population in which a proportion $p$ of voters are Democrats and $1-p$ are Republicans. Your sample size is $N=25$. Consider the random variable $S$ which is the **total** number of Democrats in your sample. What is the expected value of this random variable? Hint: it's a function of $p$.

```{r}
EofS = N*p = 25*p
```


2\. What is the standard error of $S$ ? Hint: it's a function of $p$.

```{r}
SEofS = sqrt(Np(1-p)) = sqrt(25*p(1-p))
```


3\. Consider the random variable $S/N$. This is equivalent to the sample average, which we have been denoting as $\bar{X}$. What is the expected value of the $\bar{X}$? Hint: it's a function of $p$.

```{r}
S/N = barx
EofbarX = S/25 = p
```


4\. What is the standard error of $\bar{X}$? Hint: it's a function of $p$.

```{r}
SE of barX = sqrt(p(1-p)/N)
```


5\. Write a line of code that gives you the standard error `se` for the problem above for several values of $p$, specifically for `p <- seq(0, 1, length = 100)`. Make a plot of `se` versus `p`.

```{r}
p <- seq(0, 1, length=100) # points between 0,1 at even spaces
se_p <- sqrt(p*(1-p)/25)

library(ggplot2)

dat <- data.frame(p = p, se_p = se_p)

ggplot(dat, aes(x=p, y=se_p)) +
  geom_point()
```

6\. Copy the code above and put it inside a for-loop to make the plot for $N=25$, $N=100$, and $N=1000$.

```{r}
library(gridExtra)

N <- c(25, 100, 1000)

g <- NULL #ggplot trick

for (n in 1:length(N)){
  p <- seq(0, 1, length=100) 
  se_p <- sqrt(p*(1-p)/ N[n])
  dat <- data.frame(p = p, se_p = se_p)
  g[[n]] <- ggplot(dat, aes(x=p, y=se_p)) +
    geom_point()}

grid.arrange(grobs = g, ncol=3)
```


7\. If we are interested in the difference in proportions, $p - (1-p)$, our estimate is $d = \bar{X} - (1-\bar{X})$. Use the rules we learned about sums of random variables and scaled random variables to derive the expected value of $d$.

$$p - (1-p) = 2p - 1$$
8\. What is the standard error of $d$?

$$
\sqrt{p(1-p)/N + (1-p)(1-(1-p))/N}
=
\sqrt{2p(1-p)/N}
$$
*But textbook says it is 2 x sqrt(p(1-p)/N)

9\. If the actual $p=.45$, it means the Republicans are winning by a relatively large margin since $d= -.1$, which is a 10% margin of victory. In this case, what is the standard error of $2\hat{X}-1$ if we take a sample of $N=25$?

```{r}
p = 0.45
SE = sqrt(2*p*(1-p)/25)
SE
```

10\. Given the answer to 9, which of the following best describes your strategy of using a sample size of $N=25$?

a.  The expected value of our estimate $2\bar{X}-1$ is $d$, so our prediction will be right on.
*b.  Our standard error is larger than the difference, so the chances of $2\bar{X}-1$ being positive and throwing us off were not that small. We should pick a larger sample size.*
c.  The difference is 10% and the standard error is about 0.2, therefore much smaller than the difference.
d.  Because we don't know $p$, we have no way of knowing that making $N$ larger would actually improve our standard error.
