
## Exercises

1\. Suppose you poll a population in which a proportion $p$ of voters are Democrats and $1-p$ are Republicans. Your sample size is $N=25$. Consider the random variable $S$ which is the **total** number of Democrats in your sample. What is the expected value of this random variable? Hint: it's a function of $p$.
```{r}
S=(1-p)*25
```


2\. What is the standard error of $S$ ? Hint: it's a function of $p$.
```{r}
se_S=sqrt(25)*1*sqrt(p*(1-p))
```


3\. Consider the random variable $S/N$. This is equivalent to the sample average, which we have been denoting as $\bar{X}$. What is the expected value of the $\bar{X}$? Hint: it's a function of $p$.
```{r}
S/N=((1-p)*25)/25 #=1-p
```


4\. What is the standard error of $\bar{X}$? Hint: it's a function of $p$.
```{r}
se_S/N=1*sqrt(p*(1-p)/25)
```


5\. Write a line of code that gives you the standard error `se` for the problem above for several values of $p$, specifically for `p <- seq(0, 1, length = 100)`. Make a plot of `se` versus `p`.
```{r}
p <- seq(0, 1, length = 100)
se<- sqrt(p*(1-p)/25)

dat=data.frame(p=p,se=se)
qplot(p,se)
ggplot(dat,aes(p,se))+geom_line()
```


6\. Copy the code above and put it inside a for-loop to make the plot for $N=25$, $N=100$, and $N=1000$.
```{r}
for (N in c(25,100,1000)){
  p <- seq(0, 1, length = 100)
  se<- sqrt(p*(1-p)/N)
  dat=data.frame(p=p,se=se)
  plots=ggplot(dat,aes(p,se))+geom_line()+ggtitle(N)
  print(plots)
}
```


7\. If we are interested in the difference in proportions, $p - (1-p)$, our estimate is $d = \bar{X} - (1-\bar{X})$. Use the rules we learned about sums of random variables and scaled random variables to derive the expected value of $d$.
```{r}
d=S/N-(1-S/N)  #2*S/N-1
```


8\. What is the standard error of $d$?
```{r}
se_S/N=1*sqrt(p*(1-p)/N)
se_d=2*1*sqrt(p*(1-p)/N)
```


9\. If the actual $p=.45$, it means the Republicans are winning by a relatively large margin since $d= -.1$, which is a 10% margin of victory. In this case, what is the standard error of $2\hat{X}-1$ if we take a sample of $N=25$?
```{r}
#se_d=2*1*sqrt(p*(1-p)/N)
d=0.1
2*1*sqrt(0.45*(1-0.45)/25)
```


10\. Given the answer to 9, which of the following best describes your strategy of using a sample size of $N=25$?
a.  The expected value of our estimate $2\bar{X}-1$ is $d$, so our prediction will be right on.
b.  Our standard error is larger than the difference, so the chances of $2\bar{X}-1$ being positive and throwing us off were not that small. We should pick a larger sample size.
c.  The difference is 10% and the standard error is about 0.2, therefore much smaller than the difference.
d.  Because we don't know $p$, we have no way of knowing that making $N$ larger would actually improve our standard error.

=> b
